---
title: "Aides Drug"
author: "Onalee Garusinghe"
date: "10/01/2024"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

```{r}
female<- filter(triUmp, gender=="F")
```

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=triUmp)
barchartGC(~infected + treatment,data=female, type="percent")

```

  The graph above shows a visual graphs of females only. It shows that the females who took drugs have a percentage of 60%, giving a higher chance of women getting aids than the placebo. Only 40% took the placebo.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=triUmp)
rowPerc(table1)
colPerc(table1)
```
  In the visual numerical graph results, the percentage in the group that didn't get aids is closest to 40.87%. When you look at the first two people who got aids its closest to 60%.


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

  There were three statistical tests ran, each with a probability that come from the result from the early group with females and their chances of getting aids.The first test had a probability of 0.00007686, the second 0.0001 and the third 0.00006595. The probabilities of the statistics differences between female groups and the drug/placebo tells if the drug prevents as well.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

  The graph of the results shown shows that the group of females were not helped in the prevention of aids by the use of medicine. The bar graph shows that the females given the drugs has a higher chance of getting aids. The numerical graph shows that women who took the drug were 20% more likely to get aids than those who didn't. The three statistical tests run prove this as well, as they showed that the probability of a woman getting aids on those drugs were shown.
